---
title: "Practical Machine Learning Course Project"
---
###1. Introduction
This document represents the output of the peer assessment in the practical machine coursera course. 
Because the goal of this work is to create a precdictive model, this document is basically a report describing how this model is built, how it uses cross validation, what is the expected out of sample error, and why you some choices are made.

###2. Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 

###3. Synopsis
Using data from accelerometers on the belt, forearm, arm, and dumbbell of 6 participants who were asked to perform barbell lifts correctly and incorrectly in 5 different ways, the goal of this project is to create a model that allow us to predict the manner in which they did the exercises.

###4. Init Data
```{r setoption, message=FALSE, warning=FALSE}
library(knitr)
library(caret)
library(rattle)

opts_chunk$set(cache = TRUE)

trainingURL <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testingURL <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

download.file(trainingURL, "pml-training.csv", mode = 'wb', cacheOK = TRUE )
download.file(testingURL, "pml-testing.csv", mode = 'wb', cacheOK = TRUE )

training <- read.csv("pml-training.csv", header = TRUE, stringsAsFactors = FALSE, na.strings=c("NA","#DIV/0!","") )
testing <- read.csv("pml-testing.csv", header = TRUE, stringsAsFactors = FALSE, na.strings=c("NA","#DIV/0!",""))
```
###5. Data Exploratory and Data Cleaning
```{r}
summary(training)
```
The fowllowing piace of code remove the near zero value covariats, columns filled with NAs and unneed other columns
```{r}
nzv <- nearZeroVar(training)
training <- training[,-x]
testing <- testing[,-x]

nav <- which(apply(training,2,function(x) {sum(is.na(x))}) == 0)
training <- training[,nav]
testing <- testing[,nav]
training <- training[,-c(1:6)]

training$classe <- as.factor(training$classe)
```
###6. Creates Data Sets for Cross-Refeence Validation Purposes
```{r}
set.seed(999)
inTrain <- createDataPartition( y=training$class, p=3/4, list=FALSE)
CRtraining <- training[inTrain,]
CRtesting <- training[-inTrain,]
```
###7. Model Development and Validation
###7.1 Predicting with Trees 
```{r}
modFitTrees <- train(classe ~ ., method="rpart",data=CRtraining) )
confusionMatrix(predict(modFitTrees, newdata = CRtesting), CRtesting$classe)

fancyRpartPlot(modFitTrees$finalModel)
```
###7.2 Predicting with Random Forest 
```{r}
modFitTrees <- train(classe ~ ., method="rf",data=CRtraining) )
confusionMatrix(predict(modFitRF, newdata = CRtesting), CRtesting$classe)
```

###Final Data
```{r}
confusionMatrix(predict(modFitRF, newdata = testing), testing$classe)
predict(modFitRF, newdata = testing)
```